---
title: "Results Neuromelanine sensitive MRI in SZ"
author: "Jakob A. Kaminski, Lara Wieland, Sophie Fromm"
date: "Jan 2021"
output: word_document
---

```{r setup, include=FALSE}
rm(list = ls()) 

knitr::opts_chunk$set(echo = TRUE)
# numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 1, digits = 2)
f_pvalue = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}

# install all packages if not done yet
  
if (!require(readxl)) install.packages("readxl")
if (!require(tidyverse)) install.packages("tidyverse")
if (!require(meta)) install.packages("meta")
if (!require(metafor)) install.packages("metafor")

# load all necessary packages
    
library(readxl)
library(tidyverse)
library(meta)
library(metafor)

# UNCOMMENT THE FOLLOWING TO INTEGRATE FIGURES ONCE WE MADE IT WORK FULLY
#source('/cloud/project/figures_script.R')

```

This file is created with RMarkdown and provides a reproducible results section.

# Meta-analysis

```{r meta, include=F}
#load data
#data_qmri<-read_excel("Literatur_uebersicht.xlsx") # this did not recognize comma as decimal seps
file <- "/cloud/project/data/Literatur_uebersicht.csv"
data_qmri <- read.csv(file, header = TRUE, sep = ",", quote = "\"", dec = ",", fill = TRUE,na.strings = "NA")
data_qmri[,17:29] <- sapply(data_qmri[,17:29],as.numeric)

######random and fixed effects meta-analysis
# sm = summary measure, SMD = Hedges g/Cohens d
#for random effects model with low n of studies, HKSJ method was recommended, set hksj = TRUE (https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/random.html)
resul<-data_qmri%>%
  metacont(SZn, SZmean, SZsd, HCn, HCmean, HCsd, data = ., sm="SMD", method.smd = "Hedges", studlab = Autor) #Keep method.smd = "Cohen"? because we have <20 studies, instead use "Hedges g"; LW: agreed and changed

#exclude Yamashita - did the other authors all report CR values??
resul_cr<-data_qmri%>%filter(Autor!="Yamashita")%>%
  metacont(SZn, SZmean, SZsd, HCn, HCmean, HCsd, data = ., sm="SMD", method.smd = "Hedges", studlab = Autor)
print(resul_cr)

# interestingly excluding Yamashita changes the random-effects model to significance as well

#including test for effect of medication status
##control order
# data_glu$medication_status<-factor(data_glu$medication_status, levels = c("medicated and unmedicated", "naïve", "medicated", "unclear"))

#for info on I squared
#https://wiki.joannabriggs.org/display/MANUAL/3.3.10.2+Quantification+of+the+statistical+heterogeneity%3A+I+squared

resul_metabias<-regtest(x=resul$TE, sei=resul$seTE, model="lm") # "sei" for the standard error,
print(resul_metabias)

# carries out (various versions of) Egger's regression test for funnel plot asymmetry by testing for the Y intercept = 0 from a linear regression of normalized effect estimate (estimate divided by its standard error) against precision (reciprocal of the standard error of the estimate), a significant value implicates publication bias, CAVE: less power when small number of studies included, LW: like imputation?
 
resul_trimfill<-trimfill(resul_cr)
print(resul_trimfill)
# draw a funnel plot with missing studies filled in, in our case none, the following probably should be in the figures script, just temporarily inserted for looking at it right away
funnel(resul_trimfill, legend=TRUE)

# The trim and fill method is a nonparametric (rank-based) data augmentation technique proposed by Duval and Tweedie (2000a, 2000b; see also Duval, 2005). The method can be used to estimate the number of studies missing from a meta-analysis due to the suppression of the most extreme results on one side of the funnel plot. The method then augments the observed data so that the funnel plot is more symmetric and recomputes the summary estimate based on the complete data. The trim and fill method can only be used in the context of a fixed- or random-effects model (i.e., in models without moderators). The method should not be regarded as a way of yielding a more "valid" estimate of the overall effect or outcome, but as a way of examining the sensitivity of the results to one particular selection mechanism (i.e., one particular form of publication bias).
```
The fixed effects model suggests significant differences between between patients and controls in mean estimates for neuromelanine content in the substantia nigra (SN) (d=``r resul$TE.random``[``r resul$lower.random``; ``r resul$upper.random``], z=``r resul$zval.fixed``, p=``r resul$pval.fixed``, see Figure x). Including only studies where contrast ratios of signal intensity from SN and white matter (locus ceruleus) were available, the effect remained significant (d=``r resul_cr$TE.random``[``r resul_cr$lower.random``; ``r resul_cr$upper.random``], z=``r resul_cr$zval.fixed``, p=``r resul_cr$pval.fixed``, see Figure x).

The random effects model revealed no difference between patients and controls in mean estimates (d=``r resul$TE.random``[``r resul$lower.random``; ``r resul$upper.random``], z=``r resul$zval.random``, p=``r resul$pval.random``, see Figure x). However, the group effect was rendered significant when including only studies with contrast rations available (d=``r resul_cr$TE.random``[``r resul_cr$lower.random``; ``r resul_cr$upper.random``], z=``r resul_cr$zval.random``, p=``r resul_cr$pval.random``, see Figure x)


The test for heterogeneity between studies shows a moderate amount of heterogeneity ($\tau^2$=``r resul$tau^2``, $I^2$=``r resul$I2*100``%) according to established standards.49 !!!SOPHIE ADAPTED UNTIL HERE. Formal testing for bias using funnel plot asymmetry according to Egger et al.50 revealed no significant source of bias (t=``r  resul_metabias$zval``, p=``r resul_metabias$pval``, see eFigure 5). Trim and fill method for bias adjustment suggests no missing studies (see Figure x). 

<!-- Filling in those studies, showed a slightly lower estimate in patients, but no significant overall difference in DLPFC glutamate ((d=``r resul_trimfill$TE.random``[``r resul_trimfill$lower.random``; ``r resul_trimfill$upper.random``], z=``r resul_trimfill$zval.random``, p=``r resul_trimfill$pval.random``). -->


# Subgroup Analysis NA for our sample, correct? Only a few studies collected FEP and UHR but provide no separate analyses (LW)
<!-- # Subgroup analysis -->
<!-- ```{r pressure, include=FALSE} -->
<!-- resul_by_subgroup<-data_qmri%>% -->
<!--   metacont(SZn, SZmean, SZsd, HCn, HCmean, HCsd,  data = ., sm="SMD", method.smd = "Hedges", byvar=) -->
<!-- ``` -->

<!-- As there were overall more than k=10 studies included and more than k=2 studies per subgroup, we conducted subsequent subgroup analysis.32 We found a significant between-group effect for medication status (Q=``r resul_by_subgroup$Q.b.random``, ``r f_pvalue(resul_by_subgroup$pval.Q.b.random)``). This effect was due to a significant increased glutamate level in antipsychotic naïve patients (d=``r resul_by_subgroup$TE.random.w[2]``[``r resul_by_subgroup$lower.random.w[2]``; ``r resul_by_subgroup$upper.random.w[2]``], z=``r resul_by_subgroup$zval.random.w[2]``, ``r f_pvalue(resul_by_subgroup$pval.random.w[2])`` see Figure 1). Fixed-effects results showed the same point estimate of the effect size (d=``r resul_by_subgroup$TE.fixed.w[2]``[``r resul_by_subgroup$lower.fixed.w[2]``; ``r resul_by_subgroup$upper.fixed.w[2]``], z=``r resul_by_subgroup$zval.fixed.w[2]``, ``r f_pvalue(resul_by_subgroup$pval.fixed.w[2])`` see Figure 1). Outlier detection showed no extreme data-point for this effect. -->

<!-- ```{r sepmedumed, include=FALSE} -->
<!-- data_glu_sep_med_group<-read_excel("data_glutamate_sep_med_group_09_04_2020.xlsx") -->
<!-- data_glu_sep_med_group$medication_status<-factor(data_glu_sep_med_group$medication_status, levels = c("unmedicated", "naïve", "medicated", "unclear")) -->
<!-- resul_by_subgroup<-data_glu_sep_med_group%>% -->
<!--   metacont(SCZn, SCZmean, SCZsd, HCn, HCmean, HCsd, studylab, data = ., sm="SMD", method.smd = "Cohen", byvar=medication_status) -->
<!-- ``` -->


<!-- As described in the methods section, two studies reported medicated and unmedicated subjects separately. Therefore, we repeated subgroup analysis for medication status with separation of those studies. In this analysis, we also find a significant between-group effect (Q=``r resul_by_subgroup$Q.b.random``, p=``r resul_by_subgroup$pval.Q.b.random``, eFigure2). The random and fixed-effects model showed the identical effect with increased glutamate levels in medication naïve patients compared to healthy controls. Concerning the separated studies on unmedicated patients we here find no differences in glutamate levels.  -->
<!-- ```{r patientgr, include=FALSE} -->
<!-- resul_by_subgroup<-data_glu%>% -->
<!--   metacont(SCZn, SCZmean, SCZsd, HCn, HCmean, HCsd, studylab, data = ., sm="SMD", method.smd = "Cohen", byvar=`patient group`) -->
<!-- resul_by_subgroup -->
<!-- ``` -->
<!-- Analysis of disease status showed no significant effect for studies comparing FEP vs. chronic patients (Q=``r resul_by_subgroup$Q.b.random``, p=``r resul_by_subgroup$pval.Q.b.random``, see eFigure 3). -->
<!-- ```{r metabol, include=FALSE} -->
<!-- resul_by_subgroup<-data_glu%>% -->
<!--   metacont(SCZn, SCZmean, SCZsd, HCn, HCmean, HCsd, studylab, data=., sm="SMD", method.smd = "Cohen", byvar=metabolite) -->
<!-- resul_by_subgroup -->
<!-- ``` -->
<!-- Analysis of results from different metabolite estimates (Glx vs. Glutamate) revealed no effect of subgroup (Q=``r resul_by_subgroup$Q.b.random``, p=``r resul_by_subgroup$pval.Q.b.random``, see eFigure 4). -->


# Meta-Regression

```{r metareg, include=F}

data_qmri <- data_qmri %>% rowwise() %>% mutate(age_mean=mean(c(age_HC, age_SZ), na.rm=T))

resul<-data_qmri%>%
  metacont(SZn, SZmean, SZsd, HCn, HCmean, HCsd, data = ., sm="SMD", method.smd = "Cohen")

resul_metareg_y<-metareg(resul, year)
resul_metareg_a<-metareg(resul, age_mean)
```

To control for possible further sources of variance, we conducted random-effects meta-regression. Publication year showed no significant moderating effect (QM=``r resul_metareg_y$QM``, p=``r resul_metareg_y$pval[2]``, see eFigure 7). Age of the investigated subjects did also show no significant moderating effect QM=``r resul_metareg_a$QM``, p=``r resul_metareg_a$pval[2]`` see eFigure 8).

# Meta-analysis of variance ratio

```{r meta_vr_cvr, include=F}
cvar_ratio<-data_qmri%>%
  escalc("CVR", n1i=SZn, m1i=SZmean, sd1i=SZsd, n2i=HCn, m2i=HCmean, sd2i=HCsd, data=., append = T)
resul_cvar_ratio<-rma.uni(cvar_ratio, measure = "CVR")
var_ratio<-data_qmri%>%
  escalc("VR", n1i=SZn, m1i=SZmean, sd1i=SZsd, n2i=HCn, m2i=HCmean, sd2i=HCsd, data=., append = T)
resul_var_ratio<-rma.uni(var_ratio, measure = "VR", slab=data_qmri$Autor)
```

Taking possible effects of mean differences into account, we calculated coefficient of variation ratio. The adjusted measure shows no significant difference (logCVR=``r resul_cvar_ratio$beta[1]`` [``r resul_cvar_ratio$ci.lb``;``r resul_cvar_ratio$ci.ub``]; z=``r resul_cvar_ratio$zval``; p=``r resul_cvar_ratio$pval``, see Figure x) across all studies. The calculation of a random-effects model for differences in variability ratio also revealed no significant effect in patients as compared to controls (logVR=``r resul_var_ratio$beta[1]`` [``r resul_var_ratio$ci.lb``;``r resul_var_ratio$ci.ub``]; z=``r resul_var_ratio$zval``; p=``r resul_var_ratio$pval``, see eFigure x). 

# ```{r meta_cvr_sub, include=F}
# cvar_ratio<-data_qmri%>%
#   escalc("CVR", n1i=SCZn, m1i=SCZmean, sd1i=SCZsd, n2i=HCn, m2i=HCmean, sd2i=HCsd, data=., slab=studylab, append = T)
# resul_cvar_ratio<-rma.uni(cvar_ratio, measure = "CVR", slab=data_glu$studylab, mods = ~ medication_status)
# res.n<-rma.uni(cvar_ratio, measure = "CVR", slab=data_glu$studylab, subset=(medication_status=="naïve") )
# res.m<-rma.uni(cvar_ratio, measure = "CVR", slab=data_glu$studylab, subset=(medication_status=="medicated") )
# ```

<!-- Because we found a moderating effect of medication status in mean difference meta-analysis we repeated meta-analysis of CVR and VR with medication status as moderator variable. We found a moderating effect of medication status (Q=``r resul_cvar_ratio$QM``, ``r f_pvalue(resul_cvar_ratio$QMp)``, see Figure 2) due to lower CVR in the subgroup of studies with medication naïve patients as compared to controls (logCVR=``r res.n$beta[1]``; [``r res.n$ci.lb``;``r res.n$ci.ub``]; z=``r res.n$zval``; ``r f_pvalue(res.n$pval)``). In the group of studies with medicated patients there was an increase in CVR (logCVR=``r res.m$beta[1]``; [``r res.m$ci.lb``;``r res.m$ci.ub``]; z=``r res.m$zval``; ``r f_pvalue(res.m$pval)``).   -->

# ```{r meta_vr_sub, include=F}
# var_ratio<-data_glu%>%
#   escalc("VR", n1i=SCZn, m1i=SCZmean, sd1i=SCZsd, n2i=HCn, m2i=HCmean, sd2i=HCsd, data=., slab=studylab, append = T)
# resul_var_ratio<-rma.uni(var_ratio, measure = "VR", slab=data_glu$studylab, mods = ~medication_status)
# res.m<-rma.uni(var_ratio, measure = "VR", slab=data_glu$studylab, subset=(medication_status=="medicated") )
# ```
# 
# We found a significant moderating effect of medication status on VR (Q=``r resul_var_ratio$QM``, ``r f_pvalue(resul_var_ratio$QMp)``, eFigure 11) i.e. we found higher VR in the subgroup of studies with medicated patients as compared to controls (logVR=``r res.m$beta[1]``; [``r res.m$ci.lb``;``r res.m$ci.ub``]; z=``r res.m$zval``; ``r f_pvalue(res.m$pval)``).
We additionally tested for possible effects of age and publication year on CVR and found no association (see Supplement).
